---
title: "SamKnows Data Analyst Technical Test"
author: "Blaise Alako"
date: "09/05/2022"
output: 
    html_document:
        theme: cerulean
        highlight: kate
        fig_width: 10
        fig_height: 7
        fig_caption: true
        df_print: kable
        toc: true
        toc_depth: 5
        toc_float:
          collapsed: true
          smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## 1 - Packages
### Load the libraries

```{r, warning=FALSE}
library(tidyverse)        # Core Libraries - dplyr, ggplot2
library(lubridate)        # Library for manipulating date
library(skimr)            # Library for summary statistic of the data
library(plotly)           # Library for interactive plots
library(hrbrthemes)       # A theme library
library(janitor)          # Library for cleaning data
```

## 2 - Data import and Cleaning  

Read the CSV file into a tibble
```{r}
download_speed_measurement_tbl <- read_csv('../data/download_speed_measurements.csv')
upload_speed_measurement_tbl   <- read_csv('../data/upload_speed_measurements.csv')
details_for_each_person_tbl    <- read_csv('../data/details_for_each_person.csv')
```

Exploratory analysis and Cleaning: Check for duplicate rows in the data with the **janitor** package and
generate an overview summary of the input data with the **skimr** package

Let's check that we do not have duplicate rows

```{r, results='hide'}
duplicate_download <- download_speed_measurement_tbl %>% get_dupes()
duplicate_upload   <- upload_speed_measurement_tbl %>% get_dupes()
duplicate_users    <- details_for_each_person_tbl %>%  get_dupes()
```

### Comments

* No duplicate rows found in all imported raw files


### Summary statistics

Summary statistics of the imported raw input files

#### Download speed overview

**Download_speed_mesurement_tbl** data frame broad overview.
For the purpose of simplifying **skim()** function reporting output , let's display only the variables with missing value;
other information in the comment section below was derived from the full output of the  **skim()** function
```{r}
download_speed_measurement_tbl %>%
    skimr::skim() %>% 
    filter(n_missing > 0)
```

##### Comments

* The download speed table contains **191200** rows and **4** columns
* The data is made up of three variable type (**Character, logical and numeric**)
* The Character and Logical columns **do not have missing value and looks OK**
* The numeric **measured_download_speed_in_Mbps** columns contain missing data (**~0.20%**)

#### Upload speed overview

**Upload_speed_mesurement_tbl** data frame broad overview.
For the purpose of simplifying **skim()** function reporting output , let's display only the variables with missing value;
other information in the comment section below was derived from the full output of the  **skim()** function
The code result is not shown to save space on this report and the output is summarized in the comment section below

```{r, results = 'hide'}
upload_speed_measurement_tbl %>% 
    skimr::skim()%>% 
    filter(n_missing > 0)
```

##### Comments

* The upload speed table contains **159447** rows and **4** columns
* The data is made up of three variable type (**Character, logical and numeric**)
* The Character and Logical columns **do not have missing and looks OK**
* The numeric **measured_upload_speed_in_Mbps** columns contain missing data (**~0.20%**)

#### Person details overview

**details_for_each_person_tbl** data frame broad overview.
For the purpose of simplifying **skim()** function reporting output , let's display only the variables with missing value;
other information in the comment section below was derived from the full output of the  **skim()** function
The code result is not shown to save space on this report and the output is summarized in the comment section below

```{r, results = 'hide'}
details_for_each_person_tbl %>%  
    skimr::skim() %>% 
    filter(n_missing > 0)
```

##### Comments

* The details for each person table contains **291** rows and **4** columns
* The data is made up of two variable type (**Character and numeric**)
* The Character and numeric columns **do not have missing and looks OK**


From the **DATA_DICTIONARY.md** we observe that **download_speed_measurement_tbl** and **details_for_each_person_tbl**
have a column in common, the **person_id** column
From the **DATA_DICTIONARY.md** we observe that **upload_speed_measurement_tbl** and **details_for_each_person_tbl**
have a column in common, the **person_id** column

For simplicity we will merge the data in two separate step as follows.

* 1- Merge download speed and users details data frames using the **person_id** as a common field
* 2- Merge Upload speed and users details data frames using the **person_id** as common field

```{r}

download_speed_user_detail_tbl <- download_speed_measurement_tbl %>%
    # Join by the common column (person_id)
    left_join(details_for_each_person_tbl, 
              by = c("person_id" = "person_id"))

upload_speed_user_detail_tbl <- upload_speed_measurement_tbl %>%
    # Join by person_id the common
    left_join(details_for_each_person_tbl, 
              by = c("person_id" = "person_id"))

```

#### Tasks:

We are require to filter the data on:

* **User living in specific cities**
* **Where measurement was successful**
* **Where test were conducted after a specific date**

Since we decided to process the download-user-details and upload-users_detail in two separate steps,
to avoid repeating the similar processing steps we will define some functions to
facilitate the processing.

##### Action plan: 

Define three function:

* **filter_by_city_date_test_status**         : filter the data on specified conditions
* **calculate_speed_percentile_perunit_time** : Calculate the nth percentile per defined unit time (hour, day, week, month)
* **calculate_speed_average_per_variables**   : Calculate the average download or upload speed per user

```{r}
# filter_by_city_date_test_status function
filter_by_city_date_test_status <-
    function(data    = data,                          # input data frame
             date    = '2020-12-31',                  # date: Default Date above which to select data
             cities  = c("Samsville", "Databury"),    # cities: Default cities to filter data by
             success = TRUE){                         # Success: Whether the test was a success (TRUE or FALSE)
        data <- data %>% filter(city %in% cities) %>%
            # Filter on successful test
            filter(did_test_complete_successfully == success) %>%
            # Convert the character type time_of_measurement into 
            # datetime data type using ymd_hms from the lubridate package
            mutate(time_of_measurement = ymd_hms(time_of_measurement)) %>%
            # select rows where measurement was done after a specific date
            filter(time_of_measurement > ymd(date))
        return (data)
    }

# calculate_speed_percentile_perunit_time function 
calculate_speed_percentile_perunit_time  <-
    function(data        = data,                              # Input data frame
             measurement = measured_download_speed_in_Mbps,   # unquote column name of speed measurement(download/upload)
             unit_time   = "day",                             # Unit time to compute percentile from
             probs       = .60){                              # quantile probability with value in [0,1]
        # prepare the measurement to be used inside tidy methods
        measurement_expr <- enquo(measurement)
        if (as_label(measurement_expr) %in% colnames(data) ){
            data <- data %>%  
                mutate(date = time_of_measurement %>%
                       floor_date(unit = unit_time)) %>%
                group_by(person_id, date) %>%
                # Calculate the mean upload or download speed per person per date(day)
                summarize(average_measurement = mean(!!measurement_expr)) %>%
                ungroup() %>%
                group_by(person_id) %>%
                # Calculate the nth percentile download or upload speed per user
                summarize(mesurement_percentile = quantile(average_measurement, 
                                                           probs= probs)) %>%
                ungroup()
        }
        else{
            stop(str_glue("measurement = {measurement_expr} is not a permitted option." ))
        }
        return(data)

    }
# calculate_speed_average_per_variables function
calculate_speed_average_per_variables <-
    function (data = data,                                      # input data frame
              ...,                                              # unquote variables to group by
              measurement = measured_download_speed_in_Mbps){   # unquote column name of speed measurement
        measurement_expr <- enquo(measurement)
        groupby_vars_expr <- quos(...)
        if (as_label(measurement_expr) %in% colnames(data) ){
            data <- data %>%
            group_by(!!! groupby_vars_expr) %>%
            summarise(average_measurement = mean(!!measurement_expr)) %>%
            ungroup()
        }
        else{
            stop(str_glue("measurement = {measurement_expr} is not a permitted option." ))
        }
        return(data)
}


```

Now let's combine these data frame the Tidy way into a single table which has the structure
**person_id, city, type_of_broadband_connection, name_of_isp, average_download_speed, average_upload_speed, percentile60_download_speed**

#### Subset data

Filter download and upload measurement as per instruction
```{r}

successfull_download_tbl <-
    download_speed_user_detail_tbl %>%
    # Filter data frame on given date, cities and measurement status
    filter_by_city_date_test_status(date    = '2020-12-31',
                                    cities  = c("Samsville", "Databury"),
                                    success = TRUE)

successfull_upload_tbl <- upload_speed_user_detail_tbl %>%
    filter_by_city_date_test_status(date    = '2020-12-31',
                                    cities  = c("Samsville", "Databury"),
                                    success = TRUE)
```

#####  Avg/percentile

Calculate and append average and percentiles
```{r}
download_upload_speed_users_details_tbl <- successfull_download_tbl %>%
    # Calculate the average download speed per users
    calculate_speed_average_per_variables(person_id,
                                          city,
                                          type_of_broadband_connection,
                                          name_of_isp) %>%
    # Rename the average download speed column name
    rename(average_download_speed_in_Mbps = average_measurement) %>%
    # Calculate and append the average upload speed per user
    left_join(successfull_upload_tbl %>%
                  calculate_speed_average_per_variables(
                      person_id ,
                      measurement = measured_upload_speed_in_Mbps) %>%
                  rename(average_upload_speed_in_Mbps=average_measurement)) %>%
    # Optional: Calculate and append the 60th percentile download speed per day
    left_join(successfull_download_tbl %>%
                  calculate_speed_percentile_perunit_time(
                      unit_time   = "day",
                      measurement = measured_download_speed_in_Mbps)) %>%
    # Rename the 60th percentile download speed column name
    rename(percentile60_download_speed_in_Mbps = mesurement_percentile)

```

##### Merged data overview

**download_upload_speed_users_details_tbl** data frame broad overview.
For the purpose of simplifying the **skim()** function reporting output , let's display only the variable with missing values;
other information in the comment section below was derived from the full output of the  **skim()** function
The code result is not shown to save space on this report but the output is summarized in the comment section below

```{r, results='hide'}
download_upload_speed_users_details_tbl %>%
    skimr::skim() %>% 
    filter(n_missing > 0)
```

##### Comment

* There are no missing values in the filtered and combined **download_upload_speed_users_details_tbl** data frame
as shown by the **n_missing** column of the **skim()** output
* Missing values were removed when we kept only those measurement that ran successfully see (**filter_by_city_date_test_status**) function above



## 3 - Data quality 

There are missing data in the raw input files above. However there are no missing data in the filtered and combined
data frame (download_upload_speed_users_details_tbl). see comment section above

### Download vs Upload

Relationship between download and upload speeds

```{r}
download_upload_scatter_plot  <- download_upload_speed_users_details_tbl %>%
    
    # Data manipulation: Construct a label for interactive plot
    mutate(label_text = str_glue('person_id = {person_id}
                                 download   = {round(percentile60_download_speed_in_Mbps,2)} Mbps
                                 upload     = {round(average_upload_speed_in_Mbps,2)} Mbps
                                 broadband  = {type_of_broadband_connection}
                                 ISP        = {name_of_isp}')) %>%
    mutate(type_of_broadband_connection = type_of_broadband_connection %>%
               fct_relevel("ADSL","VDSL","Fibre")) %>% 
    
    # Data Visualization
    ggplot(aes(x = percentile60_download_speed_in_Mbps, 
               y = average_upload_speed_in_Mbps,
               color = type_of_broadband_connection)) +
    geom_jitter(aes(text = label_text), width =.1 , height = .1) +
    scale_x_continuous(trans = "log10")  +
    scale_y_continuous(trans = "log10")  +
    facet_wrap(city ~ name_of_isp) + 
    labs(title    = "Average Download Upload speed by ISP by City by Connection type", 
          x       = "Average download speed in Mbps (log scale)",
          y       = "Average upload speed in Mbps (log scale)",
          color   = "Connection Type",
          caption = "B. Alako: SamKnows Data analyst Technical Test") +
     theme_ipsum()

# Interactive plot
ggplotly(download_upload_scatter_plot, tooltip = "text")

```

#### Comment
* There are mislabeled connection types having unexpected download and upload speed.
* Move the mouse over the point on the plot to get the details of mislabeled entries.

#### Action plan:

* Identify the mislabeled entries from download_upload_speed_users_details_tbl data frame 
* Relabel Connection type accordingly.


```{r}
# Identify theses mislabeled entries
fibre_labeled_as_adsl_tbl <- download_upload_speed_users_details_tbl %>% 
        filter(type_of_broadband_connection == "ADSL" ,
               percentile60_download_speed_in_Mbps > 25) %>% 
    mutate(average_download_speed_in_Mbps = round(average_download_speed_in_Mbps,2))
adsl_labeled_as_fibre_tbl <- download_upload_speed_users_details_tbl %>% 
        filter(type_of_broadband_connection == "Fibre" ,
               percentile60_download_speed_in_Mbps <  25) %>% 
     mutate(average_download_speed_in_Mbps = round(average_download_speed_in_Mbps,2))

```
##### Mislabeled Fibre
Fibre connection type labeled as ADSL connection type
```{r}
fibre_labeled_as_adsl_tbl
```
##### Mislabeled ADSL
ADSL connection type labeled as Fibre connection type
```{r}
adsl_labeled_as_fibre_tbl
```

##### Comment

* There are six ADSL connection type entries with speed in the range of Fibre speed (>180 Mbps).
* There are six Fibre connection type entries with speed in the range of ADSL speed (9-14 Mbps).

Let's correctly relabel mislabeled connection type
```{r}
download_upload_speed_users_details_tbl <- 
    download_upload_speed_users_details_tbl %>% 
    mutate(type_of_broadband_connection = case_when(
        type_of_broadband_connection == "ADSL"  & percentile60_download_speed_in_Mbps > 25 ~ "Fibre", 
        type_of_broadband_connection == "Fibre" & percentile60_download_speed_in_Mbps < 25 ~ "ADSL",  
        TRUE ~ type_of_broadband_connection
))
```


##   4 - Data Summarization
Difference in download speed between ISP for each connection type.
Since the procedure to calculate the difference in download speed and difference in upload speed between ISP for each
Connection type is the same, let's define a function to facilitate this process

```{r}
# A function to summarize the download or upload speed difference between ISP
speed_difference_btw_isp <- 
    function(data = data,                                       # input data frame
             ...,                                               # unquote variables to group by
             measurement = percentile60_download_speed_in_Mbps  # unquote column name of speed measurement
             ){
    groupby_vars_expr <- quos(...)
    measurement_expr <- enquo(measurement)
    if (as_label(measurement_expr) %in% colnames(data) ){
        data <- data %>%
            select(type_of_broadband_connection, 
                   name_of_isp, city, 
                   !! measurement_expr)  %>%
            group_by(!!! groupby_vars_expr) %>%
            summarize(mean_measurement = mean(!! measurement_expr)) %>%
            ungroup() %>%
            pivot_wider(names_from  = name_of_isp,
                        values_from = mean_measurement)
    }
    else{
        stop(str_glue("measurement = {measurement_expr} is not a permitted option." ))
    }

    return(data)
}    

```


### 4 - a) Speed difference
Download speed difference and upload speed difference between ISP for each connection type

```{r}
# Difference in Download between ISP by connection type
difference_in_download_btw_isp_tbl <- download_upload_speed_users_details_tbl %>%
    speed_difference_btw_isp(type_of_broadband_connection,
                             name_of_isp,
                             measurement = percentile60_download_speed_in_Mbps ) %>% 
    mutate(Fibrelicious = round(Fibrelicious, 2), 
           Useus = round(Useus,2))

difference_in_download_btw_isp_tbl
```

#### Comment

* Overall Fibrelicious ISP offers a better download speed compare to Useus ISP
* We will need to test for the significance of this observed difference in download speed through hypothesis testing.

```{r}

# Difference in Download between ISP by connection type by city
difference_in_download_btw_isp_by_city_tbl <- 
    download_upload_speed_users_details_tbl %>%
    speed_difference_btw_isp(type_of_broadband_connection,
                             name_of_isp,
                             city,
                             measurement = percentile60_download_speed_in_Mbps ) %>% 
    mutate(Fibrelicious = round(Fibrelicious, 2), 
           Useus = round(Useus,2))

difference_in_download_btw_isp_by_city_tbl
```

#### Comment 

* Overall Fibrelicious ISP offers a better download speed compare to Useus ISP
* We will need to test for the significance of this observed difference in download speed through hypothesis testing.
* For users living in Samsville and opting for ADSL connection type, Useus ISP offers a better download speed

```{r}

# Difference in upload speed between ISP by connection type
difference_in_upload_btw_isp_tbl <- download_upload_speed_users_details_tbl %>%
    speed_difference_btw_isp(type_of_broadband_connection,
                             name_of_isp,
                             measurement = average_upload_speed_in_Mbps ) %>% 
    mutate(Fibrelicious = round(Fibrelicious, 2), 
           Useus = round(Useus,2))

difference_in_upload_btw_isp_tbl
```

#### Comment

* Overall Fibrelicious ISP offers a better upload speed compare to Useus ISP
* We will need to test for the significance of this observed difference in upload speed through hypothesis testing.

```{r}

# Difference in Upload speed between ISP by connection type by city
difference_in_upload_btw_isp_by_city_tbl <- 
    download_upload_speed_users_details_tbl %>%
    speed_difference_btw_isp(type_of_broadband_connection,
                             name_of_isp,
                             city,
                             measurement = average_upload_speed_in_Mbps ) %>% 
    mutate(Fibrelicious = round(Fibrelicious, 2), 
           Useus = round(Useus,2))

difference_in_upload_btw_isp_by_city_tbl
```

##### Comment

* Overall Fibrelicious ISP offers a better upload speed compare to Useus ISP
* We will need to test for the significance of this observed difference in upload speed through hypothesis testing.
* For users living in Samsville and opting for ADSL connection type, Useus ISP offers a better Upload speed.



### 4 - b) Speed distribution
Distribution of average download speed by ISP by city by connection type

```{r}

# Data manipulation
download_upload_speed_distibution_plot <- 
download_upload_speed_users_details_tbl %>%
    select(type_of_broadband_connection,
           name_of_isp,
           city,
           percentile60_download_speed_in_Mbps) %>%
    mutate(type_of_broadband_connection = type_of_broadband_connection %>% 
               fct_relevel("ADSL","VDSL","Fibre")) %>% 
    # Data Visualization
    ggplot(aes(x    = percentile60_download_speed_in_Mbps, 
               fill = type_of_broadband_connection)) +
    geom_histogram(binwidth = 2, 
                   alpha    = .7, 
                   position ="identity") + 
    facet_wrap(name_of_isp~ city) +
    labs(title   = "Distribution of Average Download Speed by \nISP by City by Connection type",
         x       = "Average download speed (Mpbs)",
         y       = "Number of Users",
         fill    = "Connection Type",
         caption = "B. Alako: SamKnows Data analyst Technical Test") +
    theme_ipsum()

download_upload_speed_distibution_plot
```

##### Comment:

* Regardless of the city, the broadband download speed are within the expected range.
* There is not much difference between ADSL and VDSL download speed in Databury from Useus ISP
* Fibre download speed are higher from Fibrelicious ISP compared to Useus ISP regardless of the city
* For Users living in Databury and willing to get a Fiber connection, Fibrelicious offers a faster connection compare to Useus ISP
* For those with a limited budget in Databury, assuming fast download speed correspond to higher cost
There is no value for money buying the expensive package VDSL compared to the cheaper ADSL from Useus ISP, the download speed are almost the same.


### 4 - c) Better/Worse

Better/Worse download speed from Fibrelicious or from Useus, How much better/worse
Let's get the answer from the download-speed-between-ISP-by-city-and-by-connection-type data frame computed above

```{r}
difference_in_download_btw_isp_by_city_tbl
```

##### Comment

* A Customer living in Databury and having a Fibre connection
will have a better connection speed from the ISP provider Fibrelicious.
* The average fibre connection speed from Fibrelicious in Databury is
219.15 Mbps


### 4 - d) Speed by ISP

Let's explore the overall download speed per hour per day by ISP by connection type 
We hope to derive from this exploration an insight of the expected download difference by ISP
To achieve this let's define a function to manipulate the data frame accordingly
```{r}

# A function for data manipulation
download_speed_by_isp_per_hour <- function(data = data , isp_speed_difference = FALSE){
    data_tbl <- data %>%
    select(time_of_measurement,
           measured_download_speed_in_Mbps,
           type_of_broadband_connection,
           name_of_isp) %>%
    mutate(time_of_measurement = hour(time_of_measurement)) %>%
    group_by(type_of_broadband_connection, 
             name_of_isp, 
             time_of_measurement) %>%
    summarize(measured_download_speed_in_Mbps = mean(measured_download_speed_in_Mbps)) %>%
    ungroup() 
    if (isp_speed_difference){
        data_tbl <- data_tbl %>% 
            pivot_wider(names_from = name_of_isp, 
                        values_from = measured_download_speed_in_Mbps) %>% 
    mutate(measured_download_speed_in_Mbps = Fibrelicious - Useus) %>% 
            mutate(label_text = str_glue("download speed = {round(measured_download_speed_in_Mbps,2)} Mbps"))
    }
    else{
        data_tbl <- data_tbl %>% 
            mutate(label_text = str_glue("download speed = {round(measured_download_speed_in_Mbps,2)} Mbps"))
    }
    data_tbl <- data_tbl %>% 
    mutate(label_text = str_glue("Time: {time_of_measurement} o'clock
                                 {label_text}")) %>%
    mutate(type_of_broadband_connection = type_of_broadband_connection %>% 
               fct_relevel("ADSL","VDSL","Fibre"))
    return(data_tbl)
}

# A function for visualizing the manipulated data
plot_download_by_isp <- function(data = data, isp_speed_difference = FALSE, legend_position= "none"){
    download_speed_plot <- download_speed_by_isp_per_hour(data = data, 
                                                          isp_speed_difference = isp_speed_difference ) %>% 
        # Data visualization
        ggplot(aes(time_of_measurement,
                   measured_download_speed_in_Mbps,
                   text  = label_text,
                   color = type_of_broadband_connection,
                   group = type_of_broadband_connection)) +
        geom_point() + 
        geom_line() +
        theme_ipsum() +
        theme(legend.position = legend_position)    
    return(download_speed_plot)
} 
```

#### Overview download

Download speed by ISP by connection type per hour of the day

```{r}
download_by_isp_plot <- successfull_download_tbl %>%  
    plot_download_by_isp() +
    facet_wrap(name_of_isp~type_of_broadband_connection,
               scale = "free", ncol=3) + 
    facet_wrap(name_of_isp ~ type_of_broadband_connection,
               scale = "free", ncol=3) +
    labs(title   = "Download by ISP at different time of the day",
         x       = "Time of the day (hours)",
         y       = "Average download speed in Mbps",
         fill    = "Connection Type",
         caption = "SamKnows Data analyst assessment test")  

# Interactive plot
ggplotly(download_by_isp_plot, tooltip = "text") 
```

##### Comment

* Downloads speed are lowest between 6pm and 10pm and highest throughout 12am to 5pm
* Lowest download speed coincide with not working hours, probably when most household are connected to the net (Netflix, YouTube ... time!!)


#### Difference
Download speed difference by ISP by connection type per hour of the day

```{r}
difference_download_by_isp_plot <- successfull_download_tbl %>%  
    plot_download_by_isp(isp_speed_difference = TRUE) +
    facet_wrap(~type_of_broadband_connection,
               scale = "free", ncol=3) +
    labs(title   = "Difference in Download speed by ISP at different time of the day",
        x       = "Time of the day (hours)",
        y       = "Difference in download speed (Mbps)",
        fill    = "Connection Type",
        caption = "B. Alako: SamKnows Data analyst assessment test") 

ggplotly(difference_download_by_isp_plot, tooltip = "text")
```

##### Comment

* Downloads speed difference between ISP are lowest between 6pm and 10pm and highest throughout 12am to 5pm for both Fibre and VDSL Connection type. Corroborating the above observations.
* Overall ADSL download speed difference pattern is similar to Fibre and VDSL with lowest download speed at 3PM and 2AM


